### 2023-05-24 paper/Chapters/Chapter4.tex

\chapter{Approach}\label{cha:approach}

\section{Feature detection}\label{sec:feature_detection}

\section{Classifier}\label{sec:classifier}