### 2023-05-25 paper/Chapters/Chapter4.tex
Update the paper

\chapter{Approach}\label{cha:approach}

\section{Feature detection}\label{sec:feature_detection}

\section{Camera calibration}\label{sec:camera_calibration}

\subsection{Reprojection error}\label{sub:reprojection_error}

The reprojection error is the distance between the reprojected point and the
measured one. It is used to evaluate the quality of the camera calibration.

We chose to minimize the reprojection error between the board and back-projected
corners, which were initially detected. The projection and back-projection are
the inverse of each other, hence minimizing the error between the projection of
the board and the detected corners and minimizing the error between the
back-projection of the detected corners and the board are
equivalent\todo{Doesn't sound very rigorous to me.}.

Let's define the following variables: \(\boldsymbol{\theta} = \begin{pmatrix}
	\theta_x, \theta_y, \theta_z
\end{pmatrix}^{T}\) is the vector of Euler rotation angles, \(\mathbf{t} = \begin{pmatrix}
	t_x, t_y, t_z
\end{pmatrix}^{T}\) is the translation vector, \(\boldsymbol{\lambda} = \begin{pmatrix}
	\lambda_1, \lambda_2
\end{pmatrix}^{T}\)\todo{Justify why we used 2} is the intrinsic parameters vector, \(\) is the focal
length, and \(\mathbf{s} = \begin{pmatrix}
	s_x, s_y
\end{pmatrix}^{T}\) is the sensor size\todo{Make sure I'm using the correct terms.}.
From the input image, we know the resolution \(\mathbf{r} = \begin{pmatrix}
	r_x, r_y
\end{pmatrix}^{T}\).
From the rotation vector, we can compute the rotation matrix \(\mathbf{R}\) as:

From \(\boldsymbol{\theta}\), the rotation matrix $R$ can be calculated as follows:

\begin{equation}
	R = R(\theta_x) R(\theta_y) R(\theta_z)
\end{equation} \todo{Is the notation ok?}

where

\begin{align}
	% R_x(\phi)   & =
	% \begin{bmatrix}
	% 	1 & 0          & 0           \\
	% 	0 & \cos(\phi) & -\sin(\phi) \\
	% 	0 & \sin(\phi) & \cos(\phi)
	% \end{bmatrix} \\
	% R_y(\theta) & =
	% \begin{bmatrix}
	% 	\cos(\theta)  & 0 & \sin(\theta) \\
	% 	0             & 1 & 0            \\
	% 	-\sin(\theta) & 0 & \cos(\theta)
	% \end{bmatrix}                 \\
	% R_z(\psi)   & =
	% \begin{bmatrix}
	% 	\cos(\psi) & -\sin(\psi) & 0 \\
	% 	\sin(\psi) & \cos(\psi)  & 0 \\
	% 	0          & 0           & 1
	% \end{bmatrix}
	R(\theta_x) & =
	\begin{bmatrix}
		1 & 0              & 0               \\
		0 & \cos(\theta_x) & -\sin(\theta_x) \\
		0 & \sin(\theta_x) & \cos(\theta_x)
	\end{bmatrix} \\
	R(\theta_y) & =
	\begin{bmatrix}
		\cos(\theta_y)  & 0 & \sin(\theta_y) \\
		0               & 1 & 0              \\
		-\sin(\theta_y) & 0 & \cos(\theta_y)
	\end{bmatrix}     \\
	R(\theta_z) & =
	\begin{bmatrix}
		\cos(\theta_z) & -\sin(\theta_z) & 0 \\
		\sin(\theta_z) & \cos(\theta_z)  & 0 \\
		0              & 0               & 1
	\end{bmatrix}
\end{align}

Then, \(H\) is given by:
\begin{equation}
	H = \begin{bmatrix}
		\mathbf{r_1} & \mathbf{r_2} & \mathbf{t}
	\end{bmatrix}.
\end{equation}

We can compute the intrinsic camera matrices as follows:
\begin{equation}
	K = \begin{pmatrix}
		\frac{f r_x}{s_x} & 0                 & \frac{r_x}{2} \\
		0                 & \frac{f r_y}{s_y} & \frac{r_y}{2} \\
		0                 & 0                 & 1
	\end{pmatrix},
\end{equation}.

Then, the back-projection of a 2D point \(\mathbf{u} = \begin{pmatrix}
	u, v, 1
\end{pmatrix}\) into a scene point with \(Z = 0\) \(\mathbf{X} = \begin{pmatrix}
	X, Y, 1
\end{pmatrix}\) is given by:
\begin{equation}
	\mathbf{X} = H g_{\lambda_1, \lambda_2}(K^{-1} \mathbf{u}).
\end{equation}

\subsection{Optimization}\label{sub:optimization}

The loss function is the sum of the squared reprojection errors between the board and
the back-projected corners, which were initially detected:
\begin{equation}
	L = \sum_{i=1}^{N} \left\lVert
	H g_{\lambda_1, \lambda_2}(K^{-1} \mathbf{u_i}) -
	\mathbf{X_i} \right\rVert^2.
\end{equation}

For the initial guess, we used the randomly chosen constant small values.

The model converged to the same results compared to the initial parameters, set
using the Scaramuzza solver, unless the initial guess was very degenerate (i.e.
\(R\) was such that the board plane passed through the principal point, and all
backprojected points were projected onto the same line).

This issue also occured with random small initial values, due to the best
approximation for \(R\) which minimizes \(L\) when the distance from the
back-projected board from the measure one was high (i.e., the \(t\) was far from
the true value) being the degenerate solution\todo{Am I using the term
correctly?}. In order to avoid this issue, we first optimized only the \(t\),
until the loss function converged, meaning that \(t\) is close to the true
value.

However, another issue was that when the board was rotated close to the
\(180^{\circ}\), \(R\) once again converged to the degenerate solution. In order
to avoid this issue, we found the solution with initial \(\theta_z\) set to
value, close to \(0^{\circ}\) and \(180^{\circ}\), and then used the solution
which minimized the loss function.

\missingfigure{Show the distribution of the reprojection error}.

\section{Feature detection}\label{sec:feature_detection}

Often, not all of the board's corners were detected initially. Firstly, we
assumed that the whole board was detected, and imputed the missing points in the
board space.\todo{Add ref to the figure}. Then, we tried extending the board
points.

\missingfigure{Show the image, the respective board, and the imputed points}

We used the obtained camera parameters to then project the imputed board points
into the image space.

\section{Classifier}\label{sec:classifier}

Given the possible positions of the previously undetected corners, we used a
classifier, proposed by \cite{geigerAutomaticCameraRange2012}.
By convolving the image with the templates of the board corners\todo{I have to
reference it from the related work section}, we obtained the
probability\todo{It's not probability} of
each pixel being a corner.\todo{Also, explain in details + math how it's done}

The original paper then used a number of checks to further prune the detected
corners, and extending line by line the initial board, formed from the
random close points which made a square \todo{Make sure that's accurate, and
reference. Also, the sentence is too compicated}.
We did not require that, since we already had the board, and we directly used
the response map to find the corners.

\missingfigure{Show the map of responses, and the detected corners}

To create a training dataset, we collected the true and false positives from the
corners we already had:

For each of the detected corners on all images, we collected the values of the
response function at the previously detected corners, and around them.
We didn't collect the values of all of the pixels, because then we would have
got too optimistic values for selecting the true positives\todo{Rephrase}.

\missingfigure{Show the distribution of the response function}

We trained a binary classifier, using the collected data, and then used it to
detect the true corners.