### 2023-08-14 paper/poster.tex
Update pdfs, add poster

\documentclass[
  twocolumn,
  % 10pt,
  % english,
  % onehalfspacing,
]{article} % Method A for two-column formatting
\usepackage{preprint}

\usepackage{microtype}
\usepackage{amsmath}
\usepackage{cleveref}
\usepackage{xfrac}
\usepackage{booktabs}
\usepackage{tabularx}
\usepackage{cases}
\usepackage[noblocks]{authblk}
\usepackage[backend=biber,style=authoryear,natbib=true]{biblatex} % Use the bibtex backend with the authoryear citation style (which resembles APA)
\addbibresource{references.bib}

\newcolumntype{L}{>{\raggedright\arraybackslash}X}

% \title{Corner localization and camera calibration from imaged lattices\\
% \large supplementary material}
\title{Corner localization and camera calibration from imaged lattices
(supplementary material)}
\author{Andrii Stadnik}
\affil{Faculty of Computer Science, Ukrainian Catholic University, Lviv, Ukraine}
\author{James Pritts}
\affil{Czech Institute of Informatics, Robotics and Cybernetics at Czech Technical University in Prague, Czechia}
\date{2023}

\begin{document}

\twocolumn[
	\begin{@twocolumnfalse}
		\maketitle
	\end{@twocolumnfalse}
]

\section{Camera model}\label{sec:camera_model}

\subsection{Notation}\label{sub:notation}

\begin{table}[htbp]
	\centering
	\begin{tabularx}{\columnwidth}{lL} % use L for the long text column
		\toprule
		Term                           & Description                                                       \\
		\midrule
		\(\mathbf{u} = \begin{pmatrix}
			               u, v, 1
		               \end{pmatrix}^{T}\) & A point in the image space                                    \\
		\(\mathbf{x} = \begin{pmatrix}
			               x, y, z, 1
		               \end{pmatrix}^{T}\) & A point in the world space                                    \\
		\(R\)                          & A \(3 \times 3\) rotation matrix                                  \\
		\(\mathbf{t}\)                 & A  translation vector of length \(3\)                             \\
		\(\alpha_x, \alpha_y\)         & Scale factor in the x and y direction                             \\
		\(c_x, c_y\)                   & Coordinates of the principal point                                \\
		\(\theta\)                     & Angle between the x and y pixel axes                              \\
		\(f\)                          & Distance from the camera center to the image plane (focal length) \\
		\(f_x, f_y\)                   & Focal lengths in the x and y directions                           \\
		\(K\)                          & Intrinsic matrix                                                  \\
		\(H\)                          & A \(3 \times 3\) matrix viewing \(z=0\)                           \\
		\(\lambda_n\)                  & Distortion coefficients                                           \\
		\bottomrule
	\end{tabularx}
	\caption{Notation}
	\label{tab:notation}
\end{table}

\subsection{Extrinsic parameters}\label{sub:extrinsic_parameters}

The extrinsic parameters represent a rigid transformation from a 3-D world
coordinate system to the 3-D camera’s coordinate system.

\begin{equation*}
	\hat{\mathbf{x}} =
	R \begin{pmatrix}
		x, y, z
	\end{pmatrix}^{T} + \mathbf{t} =
	\begin{bmatrix}
		R              & \mathbf{t} \\
		\mathbf{0}^{T} & 1
	\end{bmatrix} \begin{pmatrix}
		x, y, z, 1
	\end{pmatrix}^{T},
\end{equation*}
where
\(\begin{pmatrix}
	x, y, z
\end{pmatrix}^{T}\) is a 3D scene point,
% \(R \begin{pmatrix}
% 	x, y, z
% \end{pmatrix}^{T} + \mathbf{t}\), where
\(R\) is a \(3 \times 3\) rotation matrix
and \(\mathbf{t}\) is
a \(3 \times 1\) translation vector.

Assuming coplanar scene points (\(z = 0\)):
\begin{equation*}
	\alpha \begin{pmatrix}
		u \\ v \\ 1
	\end{pmatrix} = \begin{bmatrix}
		\mathbf{r_1} & \mathbf{r_2} & \mathbf{r_3} & \mathbf{t}
	\end{bmatrix} \begin{pmatrix}
		x \\ y \\ 0 \\ 1
		% \end{pmatrix} = \begin{bmatrix}
		%   \mathbf{p_1} & \mathbf{p_2} & \mathbf{p_4}
		% \end{bmatrix} \begin{pmatrix}
		%   x \\ y \\ 1
		% \end{pmatrix}.
	\end{pmatrix} = \underbrace{\begin{bmatrix}
			\mathbf{r_1} & \mathbf{r_2} & \mathbf{t}
		\end{bmatrix}}_{H} \begin{pmatrix}
		x \\ y \\ 1
	\end{pmatrix}.
\end{equation*}

\subsection{Distortion model}
The distortion of the image is caused by the lens not being perfectly planar.
We used the division distortion model
\citep{fitzgibbonSimultaneousLinearEstimation2001} which maps a
point from a retinal plane to the
ray direction in the camera coordinates system.
\begin{equation*}
	g(\mathbf{\hat{u}}) = \begin{pmatrix}
		u, v, \psi(r(\mathbf{\hat{u}}))
	\end{pmatrix}^{T},
	\psi(\rho) = 1 + \sum_{n = 1}^{N} \lambda_n \rho^{2n},
\end{equation*}
where
\(\mathbf{\hat{u}} = \begin{pmatrix}
	u, v, 1
\end{pmatrix}^{T}\) is a point in the retinal plane
, \(r(\mathbf{\hat{u}}) = \sqrt{u^2 + v^2}\) is the radial distance from the
principal point and \(\lambda_n\) are the distortion coefficients.

\subsubsection{Back-projection using the Division Model}\label{subsub:back_projection_using_the_division_model}

The function \(\psi(\cdot)\) is not invertible in general.
Let \(\mathbf{\hat{x}} = \begin{pmatrix}
	x, y, z
\end{pmatrix}^{T} = \alpha g(\mathbf{\hat{u}})\) be a ray in the camera coordinate system.

Then,
\begin{equation} \label{eq:division_derivation1}
	\frac{\mathbf{x}}{z}  =
	\begin{pmatrix}
		\frac{x}{z}, \frac{y}{z}, 1
	\end{pmatrix}^{T}                                                                      =
	\begin{pmatrix}
		\frac{\alpha u}{\alpha \psi(r(\mathbf{\hat{u}}))},
		\frac{\alpha v}{\alpha \psi(r(\mathbf{\hat{u}}))},
		1
	\end{pmatrix}^{T} =
	\begin{pmatrix}
		\frac{u}{\psi(r(\mathbf{\hat{u}}))},
		\frac{v}{\psi(r(\mathbf{\hat{u}}))},
		1
	\end{pmatrix}^{T}.
\end{equation}

From \ref{eq:division_derivation1} we see that
\begin{equation} \label{eq:division_derivation2}
	\begin{cases}
		\frac{x}{z} = \frac{u}{\psi(r(\mathbf{\hat{u}}))} \\
		\frac{y}{z} = \frac{v}{\psi(r(\mathbf{\hat{u}}))}
	\end{cases} \implies \\
	\begin{cases}
		u = \frac{x \psi(r(\mathbf{\hat{u}}))}{z} \\
		v = \frac{y \psi(r(\mathbf{\hat{u}}))}{z}
	\end{cases}.
\end{equation}

Now, let \(\hat{r}\) be a root of
\(r(\mathbf{\hat{u}})
= \sqrt{
	\frac{x * \psi \left( \hat{r}\right)}{z}^{2} +
	\frac{y * \psi \left( \hat{r}\right)}{z}^{2}
} = \hat{r}\), assuming \(0 \leq \hat{r} \leq w\) where \(w\) is a width of
the image.

Then, \(\mathbf{\hat{u}} = f(\mathbf{x}) =
\frac{\hat{r}}{r(\mathbf{x})}\mathbf{x}\),
where \(f(\cdot)\) is the inverse of \(g(\cdot)\).

\subsection{Intrinsic parameters}
Represent a projective transformation from the 3-D camera’s coordinates into
the 2-D image coordinates.

\begin{equation*}
	K = \begin{bmatrix}
		\alpha_x & \alpha_x \cot \theta & c_x \\
		0        & \alpha_y             & c_y \\
		0        & 0                    & 1
	\end{bmatrix} \begin{bmatrix}
		f & 0 & 0 \\
		0 & f & 0 \\
		0 & 0 & 1
	\end{bmatrix} = \begin{bmatrix}
		f_x & k   & c_x \\
		0   & f_y & c_y \\
		0   & 0   & 1
	\end{bmatrix}.
\end{equation*}
For a typical camera, \(\theta = \sfrac{\pi}{2}\) and \(\alpha_x = \alpha_y\)
\citep{hartleyMultipleViewGeometry2004}:
\begin{equation*}
	K = \begin{bmatrix}
		f & 0 & c_x \\
		0 & f & c_y \\
		0 & 0 & 1
	\end{bmatrix}.
\end{equation*}

\subsection{Complete projection and backprojection}\label{sub:complete_projection_and_backprojection}

\begin{align}
	\alpha \mathbf{u} & = K f(H\mathbf{x})
	\tag{Projection} \label{eq:projection}                     \\
	\alpha \mathbf{x} & = H^{-1} g(K^{-1}\mathbf{u}) \tag{Back
		projection} \label{eq:back_projection}.
\end{align}

\section{Approach}\label{sec:approach}

\subsection{Camera calibration}\label{sub:camera_calibration}

\subsubsection{Initial approximation}\label{subsub:initial_approximation}

\subsubsection{Solving for the camera extrinsic parameters}\label{ssub:solving_for_the_camera_extrinsic_parameters}

To derive the solver for the camera extrinsic parameters, start from
\cref{eq:projection}:
\begin{align}
	\alpha \mathbf{u}                                                          & = K f(H\mathbf{x})       \\
	\alpha K^{-1} \mathbf{u}                                                   & = f(H\mathbf{x})    &  &
	\text{Move \(K\) to the left side}                                                                    \\
	\alpha g( K^{-1}\mathbf{u})                                                & = g(f(H\mathbf{x})) &  &
	\text{Set \(\mathbf{\widehat{u}} = K^{-1}\mathbf{u}\); apply \(g(\cdot)\)}                            \\
	\alpha \begin{pmatrix}
		       \widehat{u}_x \\ \widehat{u}_y \\ \psi(r(\mathbf{\widehat{u}}))
	       \end{pmatrix}^{T} & = H\mathbf{x}.
\end{align}

To eliminate the dependency on the scale \(\alpha\), multiply both sides
vectorially by \(g(\mathbf{\widehat{u}})\):
\begin{equation}
	\alpha g(\mathbf{\widehat{u}}) \times g(\mathbf{\widehat{u}})
	= g(\mathbf{\widehat{u}}) \times H\mathbf{x} \implies
	g(\mathbf{\widehat{u}})
	% \begin{pmatrix}
	% 	\widehat{u}, \widehat{v}, \psi(r(\mathbf{\widehat{u}}))
	% \end{pmatrix}^{T}
	\times \begin{bmatrix}
		\mathbf{r_1} & \mathbf{r_2} & \mathbf{t}
	\end{bmatrix} \mathbf{x} = 0
	\label{eq:eq_no_multiplier}.
\end{equation}

From \cref{eq:eq_no_multiplier}, we can see that a point contributes to three
homogeneous equations:
\begin{align}
	\widehat{v} (r_{31}x + r_{32} y + t_3) -
	g(r(\mathbf{\widehat{u}})) (r_{21}x + r_{22}y + t_2 ) & = 0
	\label{eq:scaramuzza_system_1}                              \\
	g(r(\mathbf{\widehat{u}})) (r_{11}x + r_{12}y + t_1) -
	\widehat{u} (r_{31}x + r_{32} y + t_3)                & = 0
	\label{eq:scaramuzza_system_2}                              \\
	\widehat{u} (r_{21}x + r_{22}y + t_2 ) -
	\widehat{v} (r_{11}x + r_{12}y + t_1)                 & = 0
	\label{eq:scaramuzza_system_3}.
\end{align}

Only \cref{eq:scaramuzza_system_3} is linear in the unknowns. Each point gives a
single equation. Now, by rewriting the equation in the matrix form
\(M \cdot \mathbf{h} = 0\), where
\[
	\mathbf{h} = \begin{pmatrix}
		r_{11}, r_{12}, r_{21}, r_{22}, r_{31}, r_{32}
	\end{pmatrix}^{T},
\]
we get:
\begin{equation}
	M = \begin{bmatrix}
		-\widehat{v}_1 x_1 & -\widehat{v}_1 y_1 & -\widehat{u}_1 x_1 & -\widehat{u}_1 y_1 & -\widehat{v}_1 & -\widehat{u}_1 \\
		\vdots             & \vdots             & \vdots             & \vdots             & \vdots         & \vdots         \\
		-\widehat{v}_N x_N & -\widehat{v}_N y_N & -\widehat{u}_N x_N & -\widehat{u}_N y_N & -\widehat{v}_N & -\widehat{u}_N
	\end{bmatrix}
\end{equation}, where \(N\) is the number of points.

The linear estimate of \(\mathbf{h}\) is found by minimizing \(\left\lVert M \cdot
\mathbf{h} \right\rVert ^{2}\) using SVD. The solution is known up to a scale factor.

To find \(r_{31}\) and \(r_{33}\), note that \(\mathbf{r_1}\) and
\(\mathbf{r_2}\) are orthonormal. The derivation couldn't fit here, you can find
it in the thesis paper.

\subsubsection{Extrinsic parameters}\label{subsub:extrinsic_parameters}

Now, to find the rest of the parameters, we substitute the values, found in the
previous step into \cref{eq:scaramuzza_system_1} and
\cref{eq:scaramuzza_system_2}. We assumed the number of the division model's
parameter to be equal to 2, and the scalar multiplier to be equal to 1
\cref{subsub:back_projection_using_the_division_model}:

\begin{equation}
	\begin{bmatrix}
		A_1 \rho_1^{2} & A_1 \rho_1^{4} & -v_1   \\
		C_1 \rho_1^{2} & C_1 \rho_1^{4} & -v_1   \\
		\vdots         & \vdots         & \vdots \\
		A_N \rho_N^{2} & A_N \rho_N^{4} & -v_N   \\
		C_N \rho_N^{2} & C_N \rho_N^{4} & -v_N
	\end{bmatrix} \cdot \begin{bmatrix}
		\lambda_1 \\
		\lambda_2 \\
		t_3
	\end{bmatrix} = \begin{bmatrix}
		B_1 - A_1 \\
		D_1 - C_1 \\
		\vdots    \\
		B_N - A_N \\
		D_N - C_N
	\end{bmatrix},
\end{equation}
where
\begin{align}
	A_i & = r_{21} x_i + r_{22} y_i + t_2  \\
	B_i & = v_i (r_{31} x_i + r_{32} y_i)  \\
	C_i & = r_{11} x_i + r_{12} y_i + t_1  \\
	D_i & = u_i (r_{31} x_i + r_{32} y_i).
\end{align}

The solution can be found using the least squares method.

\end{document}